Conservative Model Order Reduction for Fluid Flow

### 2018-10-05 sections/5.res_part2.tex
updated version

\begin{table} [h]
	\centering
	\caption{Geometry parameters of the quasi-1D CVRC with an oxidizer post length $L_{op}=14$ cm.}
	\centering
	\begin{tabular}{c c c c c c }
		\toprule
		\centering
		\multirow{2}{*}{Section} &
		\multicolumn{2}{c}{Oxidizer post} &
		\multirow{2}{*}{Chamber} &
		\multicolumn{2}{c}{Nozzle} \\
		\cmidrule(lr){2-3} \cmidrule(lr){5-6}
		& injector & back-step & & converging part & diverging part\\
		\midrule
		Length (cm) & 12.99 & 1.01 & 38.1 & 1.27 & 3.4 \\
		Radius (cm) & 1.02  & $1.02 \sim 2.25$ & 2.25 & $2.25 \sim 1.04$ & $1.04 \sim 1.95$ \\
		\bottomrule
		\label{tab:geometry_parameters}
	\end{tabular} 
\end{table}
The fuel is pure gaseous methane. The oxidizer is a mixture of 42\% oxygen and 58\% water (per unit mass). And is injected in the oxidizer post at a temperature $T_{ox}=1030$ K so that both water and oxygen are in the gaseous phase. The operating conditions are listed in Table \ref{tab:operating-conditions}.

\begin{table} [h]
	\centering
	\caption{CVRC operating conditions.}
	\centering
	\begin{tabular}{l l l}
		\toprule
		\centering
		Parameter & Unit & Value \\
		\midrule
		Fuel mass flow rate, $\dot{m}_{f}$ & kg/s & 0.027   \\
		Fuel temperature, $T_{f}$ & K & 300   \\
		Oxidizer mass flow rate, $\dot{m}_{ox}$ & kg/s & 0.32   \\
		Oxidizer temperature, $T_{ox}$ & K & 1030   \\
		$O_2$ mass fraction in oxidizer, $Y_{O_2}$ & -- & 42.4\%   \\
		$H_2O$ mass fraction in oxidizer, $Y_{H_2O}$ & -- & 57.6\%   \\
		Mean chamber pressure & MPa & 1.34 \\
		Equivalence ratio, $E_r$ & -- & 0.8 \\
		\bottomrule
		\label{tab:operating-conditions}
	\end{tabular} 
\end{table}
For the combustion, we consider the one-step reaction model
\begin{equation*}\label{eq:combustion}
CH_4 + 2O_2 \rightarrow CO_2 + 2H_2O
\end{equation*}
We assume that the fuel reacts instantaneously to form products, allowing us to neglect intermediate species and finite reaction rates. As the equivalence ratio is less than one, there is oxidizer left after the combustion. Therefore, only two species need to be considered: oxidizer and combustion products.

@@

Note that the right hand side in \eqref{eq:5p2.10} suggests that, in general, mass, momentum, and energy is not conserved. Furthermore, the complex coupling of the variables in \eqref{eq:5p2.3} prohibits the application of complex and implicit time integration schemes. Therefore, a quasi-skew-symmetric form, introduced in \eqref{eq:3.7}, is used for \eqref{eq:5p2.3}. It is straight-forward to check \cite{sjogreen2010skew}, for $t,s\in \mathbb R^{N}$
\begin{equation} \label{eq:5p2.9.1}
	\frac 1 2 \delta_x (st)_j + \frac 1 2 s_j \delta_x (t)_j + \frac 1 2 t_j \delta_x (s)_j = \frac 1 4 \delta_x^+( s_j + s_{j-1} )(t_j + t_{j-1}).
\end{equation}
where $\delta_x (v)_j = \left( v_{j+1} - v_{j-1} \right)/\Delta x$ is a centered finite difference approximation of the space derivative and $\delta_x^+ (v_j) = \left( v_{j+1} - v_j \right) /\Delta x$, for some $v\in \mathbb R^{N}$. Therefore,
\begin{equation} \label{eq:5p2.9.2}
	F_{i+1/2}^{\Delta}(s_jt_j,s_{j+1}t_{j+1}) = ( s_j + s_{j-1} )(t_j + t_{j-1}),
\end{equation}
can be interpreted as an approximation of a quadratic flux function at the boundary of two adjacent finite volume cells. A better approximation of the flux in \eqref{eq:5p2.9.2} corresponds to a higher order skew-symmetric form for a quadratic variable $st$ in \eqref{eq:5p2.9.1}. We discretize the real line into N uniform cells of size $\Delta x$. A quasi-skew-symmetric form for \eqref{eq:5p2.10} now takes the form
\begin{equation}
\begin{aligned}
	\frac{d}{dt} q^i_j + \delta^+ F^{\Delta}_{i+1/2}(q^i_jr^i_j,q^i_{j+1}r^i_{j+1}) &- \delta^+ F_d^{\Delta}(d^i_j,d^i_{j+1}) + \delta^+ F_p^{\Delta}(p_j,p_{j+1}) \\
	&= \int_{c_j} s_A + s_f + s_q \ dx.
\end{aligned}
\end{equation}
for $j=1,\dots,N$. Here, $c_j$ is the $j$th cell, $q^i_j = \int_{c_j} v^i \ dx$ is the cell average of the $i$th component of $v$, $F^{\Delta}_p$ is the flux approximation of the pressure term, $F^{\Delta}_d$ is the flux approximation for the viscous term and $r = ( u , u , u , u )^T$.

The three-stage Runge-Kutta (SSP RK3) \cite{jiang1996efficient} is used to integrate \eqref{eq:5p2.10} in time.
The pressure profile for the steady state, with $q'=0$, and the pressure oscillatory mode in the unsteady phase is presented in \Cref{fig:5p2.2a}.

@@

The discontinuities that appear in the solution of \eqref{eq:5p2.10} suggests that a relatively large basis is required to resolve fine structures in the solution. Here, a POD basis is generated with $k=200$, $k=300$ and $k=400$ number of basis vectors. To avoid basis changes in the reduced system, only one POD basis is considered for $\rho$, $\rho u$ $\rho E$ and $\rho Y_{ox}$. The explicit SSP RK3 is then used to integrated the reduced system in time, for the unsteady system. The source terms are evaluated in the high-fidelity space and projected onto the reduced space. However, in principle, the DEIM can be applied to accelerate this component.

\Cref{fig:5p2.2c} shows the approximation error of the pressure, due to MOR. It is observed that the approximation is consistently improved as the number of basis vectors increases. Furthermore, the approximate solution maintains high accuracy over a relatively long time-integration. The oscillations of pressure is demostrated in \Cref{fig:5p2.2d}. The overall behaviour of pressure is well approximated by the reduced system. Similar results are obtained for a POD basis with higher number of modes.

>>>>

\begin{table} [h]
	\centering
	\caption{Geometry parameters of the quasi-1D CVRC with an oxidizer post length $L_{op}=14$ cm.}
	\centering
	\begin{tabular}{c c c c c c }
		\toprule
		\centering
		\multirow{2}{*}{Section} &
		\multicolumn{2}{c}{Oxidizer post} &
		\multirow{2}{*}{Chamber} &
		\multicolumn{2}{c}{Nozzle} \\
		\cmidrule(lr){2-3} \cmidrule(lr){5-6}
		& injector & back-step & & converging part & diverging part\\
		\midrule
		Length (cm) & 12.99 & 1.01 & 38.1 & 1.27 & 3.4 \\
		Radius (cm) & 1.02  & $1.02 \sim 2.25$ & 2.25 & $2.25 \sim 1.04$ & $1.04 \sim 1.95$ \\
		\bottomrule
		\label{tab:geometry_parameters}
	\end{tabular} 
\end{table}
The fuel is pure gaseous methane. The oxidizer is a mixture of 42\% oxygen and 58\% water (per unit mass), and is injected in the oxidizer post at a temperature $T_{ox}=1030$ K so that both water and oxygen are in the gaseous phase. The operating conditions are listed in Table \ref{tab:operating-conditions}.

\begin{table} [h]
	\centering
	\caption{CVRC operating conditions.}
	\centering
	\begin{tabular}{l l l}
		\toprule
		\centering
		Parameter & Unit & Value \\
		\midrule
		Fuel mass flow rate, $\dot{m}_{f}$ & kg/s & 0.027   \\
		Fuel temperature, $T_{f}$ & K & 300   \\
		Oxidizer mass flow rate, $\dot{m}_{ox}$ & kg/s & 0.32   \\
		Oxidizer temperature, $T_{ox}$ & K & 1030   \\
		$O_2$ mass fraction in oxidizer, $Y_{O_2}$ & -- & 42.4\%   \\
		$H_2O$ mass fraction in oxidizer, $Y_{H_2O}$ & -- & 57.6\%   \\
		Mean chamber pressure & MPa & 1.34 \\
		Equivalence ratio, $E_r$ & -- & 0.8 \\
		\bottomrule
		\label{tab:operating-conditions}
	\end{tabular} 
\end{table}
For the combustion, we consider the one-step reaction model
\begin{equation*}\label{eq:combustion}
CH_4 + 2O_2 \rightarrow CO_2 + 2H_2O.
\end{equation*}
We assume that the fuel reacts instantaneously to form products, allowing us to neglect intermediate species and finite reaction rates. As the equivalence ratio is less than one, there is oxidizer left after the combustion. Therefore, only two species need to be considered: oxidizer and combustion products.

@@

Note that the right hand side in \eqref{eq:5p2.10} suggests that, in general, mass, momentum, and energy is not conserved. Furthermore, the complex coupling of the variables in \eqref{eq:5p2.3} and the non-constant adiabatic gas index prohibits the application of complex and implicit time integration schemes. Therefore, a quasi-skew-symmetric form, introduced in \eqref{eq:3.7}, is used for \eqref{eq:5p2.3}. It is straight-forward to check \cite{sjogreen2010skew}, for $t,s\in \mathbb R^{N}$
\begin{equation} \label{eq:5p2.9.1}
	\frac 1 2 \delta_x (st)_j + \frac 1 2 s_j \delta_x (t)_j + \frac 1 2 t_j \delta_x (s)_j = \frac 1 4 \delta_x^+( s_j + s_{j-1} )(t_j + t_{j-1}).
\end{equation}
where $\delta_x (v)_j = \left( v_{j+1} - v_{j-1} \right)/\Delta x$ is centered finite difference approximation of the space derivative and $\delta_x^+ (v_j) = \left( v_{j+1} - v_j \right) /\Delta x$, for some $v\in \mathbb R^{N}$. Therefore,
\begin{equation} \label{eq:5p2.9.2}
	F_{i+1/2}^{\Delta}(s_jt_j,s_{j+1}t_{j+1}) = ( s_j + s_{j-1} )(t_j + t_{j-1}),
\end{equation}
can be interpreted as an approximation of a quadratic flux function at the boundary of two adjacent finite volume cells. A better approximation of the flux in \eqref{eq:5p2.9.2} corresponds to a higher order skew-symmetric form for a quadratic variable $st$ in \eqref{eq:5p2.9.1}. We discretize the real line into N uniform cells of size $\Delta x$. A quasi-skew-symmetric form for \eqref{eq:5p2.10} now takes the form
\begin{equation}
\begin{aligned}
	\frac{d}{dt} q^i_j + \delta^+ F^{\Delta}_{i+1/2}(q^i_jr^i_j,q^i_{j+1}r^i_{j+1}) &- \delta^+ F_d^{\Delta}(d^i_j,d^i_{j+1}) + \delta^+ F_p^{\Delta}(p_j,p_{j+1}) \\
	&= \int_{c_j} s_A + s_f + s_q \ dx.
\end{aligned}
\end{equation}
for $j=1,\dots,N$. Here, $c_j$ is the $j$th cell, $q^i_j = \int_{c_j} v^i \ dx$ is the cell average of the $i$th component of $v$, $F^{\Delta}_p$ is the flux approximation of the pressure term, $F^{\Delta}_d$ is the flux approximation for the viscous term and $r = ( u , u , u , u )^T$.

The three-stage Runge-Kutta (SSP RK3) \cite{jiang1996efficient} is used to integrate \eqref{eq:5p2.10} in time.
The pressure profile for the steady state, with $q'=0$, and the pressure oscillatory mode in the unsteady phase is presented in \Cref{fig:5p2.2a,fig:5p2.2b}, respectively.

@@

The discontinuities that appear in the solution of \eqref{eq:5p2.10} suggests that a relatively large basis is required to resolve fine structures in the solution. Here, a POD basis is generated with $k=200$, $k=300$ and $k=400$ number of basis vectors. To avoid basis changes in the reduced system, only one POD basis is considered for $\rho$, $\rho u$ $\rho E$ and $\rho Y_{ox}$. The explicit SSP RK3 is then used to integrated the reduced system in time, for the unsteady system. The source terms are evaluated in the high-fidelity space and projected onto the reduced space. However, in principle, the DEIM can be applied to accelerate the evaluation this component.

\Cref{fig:5p2.2c} shows the approximation error of the pressure, due to MOR. It is observed that the approximation is consistently improved as the number of basis vectors increases. Furthermore, the approximate solution maintains high accuracy over a relatively long time-integration. The oscillations of pressure is demonstrated in \Cref{fig:5p2.2d}. The overall behaviour of pressure is well approximated by the reduced system. Similar results are obtained for a POD basis with higher number of modes.
